Evaluating the Performance and Energy Efficiency of the \textsc{COSMO-ART} model system

### 2014-04-25 trunk/paper_enahpc14/Sec3:Environment-setup.tex
Manuel's changes

\begin{itemize}
\item \monch is a 10 rack NEC-provided cluster composed
  of 312 compute  nodes %and utilized by members  of the Swiss Platform
  %for Advanced Scientific  Computing (PASC) project~\citep{PASC}
   at CSCS. Each
  node comprises two Intel Ivy Bridge EP  E5-2660 ten-core processors
  operating at  2.2 GHz, equipped with 32\,GB of DDR3 1600\,MHz  RAM and connected  via InfiniBand Mellanox  SX6036 and FDR switches. For our  experiments, 
  a full rack constituted of 52 nodes was used. \monch is  slated to stay
  in  service  without  hardware  upgrade  for  the  duration  of  the
  Exa2Green  project,  allowing   an  identical  configuration  to  be
  recreated for future assessments of the baseline.\\

\item \pilat is a dual-socket eight-core Intel Sandy
  Bridge EP based cluster composed of 42 compute nodes %and used as Piz Daint pre-post  processing cluster 
  at  CSCS.  The 2 login  nodes and
  the 42 computes nodes consists in 11 twin-pair Intel E5-Series DALCO
  r2264i4t 2U scalable compute modules. Each module contains 4 compute
  nodes based  on two Intel Xeon E5-2670 processors operating  at 2.6\,GHz equipped with 64GB of DDR3 1600\,MHz RAM and connected  via InfiniBand Mellanox SX6036 and FDR switches. 
  For our  experiments, a  full rack constituted  of 42 standard  compute nodes
  was used. \pilat  is based  on  Intel's second-generation
  processors, conventional in  HPC systems and known to  be more power
  consuming than Ivy Bridge successor.\\

@@

We present here the two different  frameworks deployed on the testbed HPC systems to
measure  performance and  power  consumption of the baseline execution of \cosmoart

@@

\cosmoart  is compiled  using the  Extrae  compiler wrappers,
which automatically  instrument the Fortran code of  the model.  Next,
\cosmoart  is  run on  the  nodes,  thus dissipating  certain
amount  of power.   These  nodes are connected to power distribution units (PDUs)
APC8653 that account for the dissipated power with a sampling frequency of 1\,Hz. The power data captured of these PDUs is afterwards collected by the tracing server. The client,  meanwhile, sends start/stop  primitives  in  order  to  gather  captured  data  by  the watt-meters onto the  tracing server, where an instance  of the \pmlib server is running.

Once  the run  is finished,  a file  containing the  power  profile is
created using the power data  received from the tracing server and the
instrumentation post-processing generates the performance trace files.
All  these files  are  combined then  in  Paraver which  allows us  to
visualise   the   performance  trace   and   the   power  profile   of
\cosmoart all together.

>>>>

\begin{itemize}
\item \monch is a 10 rack NEC-provided cluste rat CSCS composed
  of 312 compute  nodes. %and utilized by members  of the Swiss Platform
  %for Advanced Scientific  Computing (PASC) project~\citep{PASC}
  Each node comprises two Intel Ivy Bridge EP  E5-2660 ten-core processors
  operating at  2.2 GHz, equipped with 32\,GB of DDR3 1600\,MHz  RAM and connected  via InfiniBand Mellanox  SX6036 and FDR switches. For our  experiments, 
  a full rack constituted of 52 nodes was used. \monch is  slated to stay
  in  service  without  hardware  upgrade  for  the  duration  of  the
  Exa2Green  project,  allowing   an  identical  configuration  to  be
  recreated for future assessments of the baseline.\\

\item \pilat is a dual-socket eight-core Intel Sandy
  Bridge EP based cluster at  CSCS composed of 42 compute nodes. %and used as Piz Daint pre-post  processing cluster 
  The 2 login  nodes and
  the 42 computes nodes consists in 11 twin-pair Intel E5-Series DALCO
  r2264i4t 2U scalable compute modules. Each module contains 4 compute
  nodes based  on two Intel Xeon E5-2670 processors operating  at 2.6\,GHz equipped with 64GB of DDR3 1600\,MHz RAM and connected  via InfiniBand Mellanox SX6036 and FDR switches. 
  For our  experiments, a  full rack constituted  of 42 standard  compute nodes
  was used. \pilat  is based  on  Intel's second-generation
  processors, conventional in  HPC systems and known to  be more power
  consuming than Ivy Bridge successor.\\

@@

We present here the two different  frameworks deployed on the testbed HPC systems to
measure  performance and  power  consumption of the baseline execution of \cosmoart.

@@

\cosmoart  is compiled and linked with the Extrae instrumentation library,
which automatically  instrument the Fortran~90 code of the model, e.g., to record MPI calls.  Next, \cosmoart  is  run on  the  nodes,  thus dissipating  certain
amount  of power.   These  nodes are connected to power distribution units (PDUs)
APC 8653 that account for the dissipated power with a sampling frequency of 1\,Hz. The power data captured by these PDUs is collected into the tracing server. The client,  meanwhile, sends a start/stop  primitive  in  order  to  gather  captured  data  by  the PDUs.

Once  the run  is finished,  a power-performance trace file containing the power 
data received from the tracing server and the instrumentation post-process is generated.
This file is finally loaded in Paraver in order to visualise the performance trace and to correlate it with the power profile generated by \cosmoart.
